\pkg{NeuralNetTools}: Visualization and Analysis Tools for Neural Networks

### 2015-09-28 nnt_manu.Rnw
plot section done

<<eval = T, echo= F, message = F>>=
library(NeuralNetTools)
@
<<results = 'hide'>>=
# set random seed for initial inputs during training
set.seed(123)

@@

The \pkg{RSNNS} package provides several algorithms that can be used to prune connections or nodes in a neural network.  This approach can remove connection weights between layers or input nodes that do not contribute to the predictive performance of the network.  Pruning has the benefit of a potential increase in power by increasing the degrees of freedom or the identification of input nodes with low importance.  Algorithms in \pkg{RSNNS} for weight pruning include magnitude based pruning, Optimal Brain Damage, and Optimal Brain Surgeon, whereas algorithms for node pruning include skeletonization and the non-contributing units method \citep{Zell98}.  The \code{plotnet} function includes options to display pruned connections that are removed using any of the three methods for removing weighted connections.

<<plotprune, eval = F, echo = T, fig.height = 7, fig.width = 7, out.width = '\\textwidth'>>=

# pruned model using code from RSSNS pruning demo
pruneFuncParams <- list(max_pr_error_increase = 10.0, pr_accepted_error = 1.0,
 no_of_pr_retrain_cycles = 1000, min_error_to_stop = 0.01, init_matrix_value = 1e-6,
 input_pruning = TRUE, hidden_pruning = TRUE)
mod <- mlp(x, y, size = 5, pruneFunc = "OptimalBrainSurgeon",
 pruneFuncParams = pruneFuncParams)

plotnet(mod)
plotnet(mod, prune_col = 'lightblue')
@

@@

>>>>

<<eval = T, echo= F, message = F>>=
library(NeuralNetTools)
@
<<results = 'hide'>>=
# set random seed for initial weights for training
set.seed(123)

@@

The \pkg{RSNNS} package provides several algorithms that can be used to prune connections or nodes in a neural network.  This approach can remove connection weights between layers or input nodes that do not contribute to the predictive performance of the network.  Pruning has the benefit of a potential increase in power by increasing the degrees of freedom or the identification of input nodes with low importance.  Algorithms in \pkg{RSNNS} for weight pruning include magnitude based pruning, Optimal Brain Damage, and Optimal Brain Surgeon, whereas algorithms for node pruning include skeletonization and the non-contributing units method \citep{Zell98}.  The \code{plotnet} function can be used to plot a pruned neural network, with options to omit or display the pruned connections (\cref{fig:plotprune}).

<<plotprune, eval = T, echo = T, fig.show = 'hold', fig.height = 7, fig.width = 7.7, out.width = '0.48\\textwidth', fig.cap = 'A pruned neural network using the ``Optimal Brain Surgeon\'\' algorithm described in \\citet{Zell98}.  The default plotting behavior of \\code{plotnet} is to omit pruned connections (left), whereas they can be viewed as dashed lines by including the \\code{prune\\_col} argument (right).'>>=
# pruned model using RSNNS
pruneFuncParams <- list(max_pr_error_increase = 10.0, pr_accepted_error = 1.0,
  no_of_pr_retrain_cycles = 1000, min_error_to_stop = 0.01, 
  init_matrix_value = 1e-6, input_pruning = TRUE, hidden_pruning = TRUE)
mod <- mlp(x, y, size = 5, pruneFunc = "OptimalBrainSurgeon",
 pruneFuncParams = pruneFuncParams)

# plot, pruned connections are omitted (default) or included with prune_col
par(mar = c(0, 0, 0, 0))
plotnet(mod)
plotnet(mod, prune_col = 'lightblue')
@

@@

The primary benefit of visualizing a neural network interpretation diagram is the ability to evaluate network archicture and the variation in weighted connections between the layers.  For most networks,

### 2015-09-28 nnt_manu.tex
plot section done

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# set random seed for initial inputs during training}
\hlkwd{set.seed}\hlstd{(}\hlnum{123}\hlstd{)}

@@

The \pkg{RSNNS} package provides several algorithms that can be used to prune connections or nodes in a neural network.  This approach can remove connection weights between layers or input nodes that do not contribute to the predictive performance of the network.  Pruning has the benefit of a potential increase in power by increasing the degrees of freedom or the identification of input nodes with low importance.  Algorithms in \pkg{RSNNS} for weight pruning include magnitude based pruning, Optimal Brain Damage, and Optimal Brain Surgeon, whereas algorithms for node pruning include skeletonization and the non-contributing units method \citep{Zell98}.  The \code{plotnet} function includes options to display pruned connections that are removed using any of the three methods for removing weighted connections.

\begin{kframe}
\begin{alltt}
\hlcom{# pruned model using code from RSSNS pruning demo}
\hlstd{pruneFuncParams} \hlkwb{<-} \hlkwd{list}\hlstd{(}\hlkwc{max_pr_error_increase} \hlstd{=} \hlnum{10.0}\hlstd{,} \hlkwc{pr_accepted_error} \hlstd{=} \hlnum{1.0}\hlstd{,}
 \hlkwc{no_of_pr_retrain_cycles} \hlstd{=} \hlnum{1000}\hlstd{,} \hlkwc{min_error_to_stop} \hlstd{=} \hlnum{0.01}\hlstd{,} \hlkwc{init_matrix_value} \hlstd{=} \hlnum{1e-6}\hlstd{,}
 \hlkwc{input_pruning} \hlstd{=} \hlnum{TRUE}\hlstd{,} \hlkwc{hidden_pruning} \hlstd{=} \hlnum{TRUE}\hlstd{)}
\hlstd{mod} \hlkwb{<-} \hlkwd{mlp}\hlstd{(x, y,} \hlkwc{size} \hlstd{=} \hlnum{5}\hlstd{,} \hlkwc{pruneFunc} \hlstd{=} \hlstr{"OptimalBrainSurgeon"}\hlstd{,}
 \hlkwc{pruneFuncParams} \hlstd{= pruneFuncParams)}

\hlkwd{plotnet}\hlstd{(mod)}
\hlkwd{plotnet}\hlstd{(mod,} \hlkwc{prune_col} \hlstd{=} \hlstr{'lightblue'}\hlstd{)}
\end{alltt}
\end{kframe}

>>>>

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# set random seed for initial weights for training}
\hlkwd{set.seed}\hlstd{(}\hlnum{123}\hlstd{)}

@@

The \pkg{RSNNS} package provides several algorithms that can be used to prune connections or nodes in a neural network.  This approach can remove connection weights between layers or input nodes that do not contribute to the predictive performance of the network.  Pruning has the benefit of a potential increase in power by increasing the degrees of freedom or the identification of input nodes with low importance.  Algorithms in \pkg{RSNNS} for weight pruning include magnitude based pruning, Optimal Brain Damage, and Optimal Brain Surgeon, whereas algorithms for node pruning include skeletonization and the non-contributing units method \citep{Zell98}.  The \code{plotnet} function can be used to plot a pruned neural network, with options to omit or display the pruned connections (\cref{fig:plotprune}).

\begin{kframe}
\begin{alltt}
\hlcom{# pruned model using RSNNS}
\hlstd{pruneFuncParams} \hlkwb{<-} \hlkwd{list}\hlstd{(}\hlkwc{max_pr_error_increase} \hlstd{=} \hlnum{10.0}\hlstd{,} \hlkwc{pr_accepted_error} \hlstd{=} \hlnum{1.0}\hlstd{,}
  \hlkwc{no_of_pr_retrain_cycles} \hlstd{=} \hlnum{1000}\hlstd{,} \hlkwc{min_error_to_stop} \hlstd{=} \hlnum{0.01}\hlstd{,}
  \hlkwc{init_matrix_value} \hlstd{=} \hlnum{1e-6}\hlstd{,} \hlkwc{input_pruning} \hlstd{=} \hlnum{TRUE}\hlstd{,} \hlkwc{hidden_pruning} \hlstd{=} \hlnum{TRUE}\hlstd{)}
\hlstd{mod} \hlkwb{<-} \hlkwd{mlp}\hlstd{(x, y,} \hlkwc{size} \hlstd{=} \hlnum{5}\hlstd{,} \hlkwc{pruneFunc} \hlstd{=} \hlstr{"OptimalBrainSurgeon"}\hlstd{,}
 \hlkwc{pruneFuncParams} \hlstd{= pruneFuncParams)}

\hlcom{# plot, pruned connections are omitted (default) or included with prune_col}
\hlkwd{par}\hlstd{(}\hlkwc{mar} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{,} \hlnum{0}\hlstd{,} \hlnum{0}\hlstd{,} \hlnum{0}\hlstd{))}
\hlkwd{plotnet}\hlstd{(mod)}
\hlkwd{plotnet}\hlstd{(mod,} \hlkwc{prune_col} \hlstd{=} \hlstr{'lightblue'}\hlstd{)}
\end{alltt}
\end{kframe}\begin{figure}[!ht]

{\centering \includegraphics[width=0.48\textwidth]{figs/plotprune-1} 
\includegraphics[width=0.48\textwidth]{figs/plotprune-2}

\caption{A pruned neural network using the ``Optimal Brain Surgeon'' algorithm described in \citet{Zell98}.  The default plotting behavior of \code{plotnet} is to omit pruned connections (left), whereas they can be viewed as dashed lines by including the \code{prune\_col} argument (right).}\label{fig:plotprune}
\end{figure}
